---
title: "Toronto Bike Parking Map (February 2021)"
author: "Laura Cline"
date: "23/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The `leaflet` package is originally a JavaScript library of the same name that has been brought over to R. It makes it easy to make interactive maps. The basics are fairly similiar to the `ggmap` set-up, but there are many other new options. 

In the same way as a graph in `ggplot` function, a map in the leaflet package begins with a call to the `leaflet()` function. This allows you to specify data, and a bunch of other options such as width and height. After this, we add 'layers' in the same way that we added them in `ggplot`. The first layer that we will add is a tile with the function `addTiles()`. In this case, the default is from OpenStreetMap. After that, we will add markets that show the location of each bike parking spot with `addMarkers()`. 

## Load and Clean Data
```{r}
# This code is based on code from: https://open.toronto.ca/dataset/street-furniture-bicycle-parking/.
# install.packages("opendatatoronto")

library(opendatatoronto)
library(tidyverse)

raw_data <- read_csv("bike_parking.csv")
raw_data
```

```{r}
# Remove irrelevant dimensions 

bike_data <- tibble(ward = raw_data$WARD,
                    id = raw_data$ID,
                    status = raw_data$STATUS,
                    street_address = paste(raw_data$ADDRESSNUMBERTEXT, raw_data$ADDRESSSTREET),
                    latitude = raw_data$LATITUDE,
                    longitude = raw_data$LONGITUDE)
```

```{r}
# Remove the temporary bike racks 

bike_data <-
  bike_data %>%
  filter(status == "Existing") %>%
  select(-status)

bike_data <-
  bike_data %>%
  filter(ward == 11) %>%
  select(-ward)
```

```{r}
#Write dimension reduced data to a csv 

write_csv(bike_data, "bike_data.csv")
```

# Static Map of Bike Parking Locations in Toronto (February 2021)


# Interactive Map of Bike Parking Locations in Toronto (February 2021)

```{r}
# install.packages("leaflet")
library(leaflet)

leaflet(data = bike_data) %>%
  addTiles() %>%
  addMarkers(lng = bike_data$longitude,
             lat = bike_data$latitude,
             popup = bike_data$street_address)
```

